\cleardoublepage
\chapter{Geometric Disk Cover}
\label{chap:approx}
\lhead{Chapter \ref{chap:approx}. \emph{\nameref{chap:approx}}} % This is for the header

The algorithms in the previous chapter have some drawbacks when used in a web application. Not only were the running times too large to be used in real-time, but they also required some a priori knowledge about the number of clusters on a given window, which is infeasible, since there is no efficient way to infer how many clusters there are in a new region, without testing for all possible values of $K$. This approach should be able to calculate the final number of selected points, constrained to a minimum distance factor between them.

Given a set $N$ of $n$ points and a minimum distance $d$, find a set of centroids $C$ of size $k$, such that every point in $N$ will covered by a disk with radius $d$ and centre in one of the points in $C$ whilst minimising the value of $k$. This problem is known as \emph{geometric disk cover}\cite{geodisk}. This chapter describes an efficient way to approximate the optimal solution to this new problem. The calculated solution will be bound above by $m \ln {n}$ points, where $m$ is the optimal value of $k$.

\section{Approximation Algorithm}
Calculating the optimal solution for Geometric Disk Cover is {NP-hard} \ref{diskcomplex}. However, there are ways of finding a reasonable approximation to the optimal solution in a short amount of time. Briefly, this approach requires two steps. The first step is the Proximity Graph Building step, which builds a graph connecting all pairs of points that are close together. The second step is a Set Cover step, which uses an approximation algorithm to cover the graph built in the former step. The following sections describe these steps in more detail.

\subsubsection*{Proximity Graph Construction}
The first step is to create a graph connecting all pairs points that are within a distance of each other. This means constructing a graph in which every point is a vertex, and every edge connects two vertices whose points are within the given distance, also known as a proximity graph \cite{proximity}. The naïve approach to building this graph would be to test all the distances between each pair of points, taking $\bigo(n^2)$ operations. Alternatively a line sweep algorithm or a series of range searches on a \kdtree could be used to speed up the process. The line sweep algorithm would require a $\bigo(n \log{n})$ sorting algorithm, followed by $\bigo(n^2)$ comparisons. However, the latter is limited by the number of points within the sliding window, which should only contain a fraction of the total number of points on the map, since the distance chosen is a fraction of its dimensions. The \kdtree range search, on the other hand, requires a $\bigo(n \log{n})$ construction of a \kdtree using a median of medians algorithm. This is followed by $n$ queries consisting of finding the points within a square of side $2d$ centred around each of the points. Each one of these operations take $\bigo(\sqrt{N})$ time \cite{kdrange}. Even though the complexity of the \kdtree range search is theoretically faster than the line sweep method, it comes with an extra overhead of handling a more complex structure. As such, both methods are analysed from an experimental point of view in this thesis. After this operation, the graph connecting all neighbours is built. Figure \ref{fig:aa1} illustrates one such graph:

\input{Figures/apx1}

This graph is represented via adjacency lists. The adjacency lists have an advantage over an adjacency matrix as linked lists can be used to considerably reduce the memory footprint in sparse graphs. It also allows for faster sequential access to the neighbours of any given point. These linked lists share similarities to the half-edge structure, in which each node represents a unidirectional edge that contains a pointer to its counterpart.

\subsubsection*{Set Cover}
The second step to the algorithm is to chose a small subset of vertices whose neighbours unions are equal to the whole set of graph's vertices. This is known as \emph{set cover}, and its solution can be approximated using a greedy approach \cite{greedyapprox}. Starting with the whole uncovered graph, the point with the largest number of uncovered neighbours is selected. This is done iteratively until no uncovered points remain on the graph. This approach approximates the number of points to within $m \ln{n}$ points, were $m$ is the optimal number.

At each step of this algorithm, one point $p$ and all its neighbours and respective edges must be removed from the graph. This is done by iterating through the adjacency lists of the neighbours of $p$ and deleting all the connections to their neighbours (second-degree neighbours of $p$). This operation takes exactly $\bigo(n)$ time where $n$ is proportional to the number of edges to be deleted. Figure \ref{fig:aa2} illustrates the graph after the first iteration:

\input{Figures/apx2}

After all the points are covered, the centers make the subset of centroids that will be displayed as the final output. The cardianality of this set will not exceed the approximation as described above.

\input{Figures/apx3}
\subsection{Results}
This section reports both approaches (\kdtrees and line sweeping) to the graph building portion of the algorithm. The value of $d$ is the radius of the disks. It is given as a percentage of the largest dimension of the window, as this value is easier to adjust and analyse by a human user. Figure \ref{fig:sweden_dist} shows the outputs for different values of $d$.

\input{Figures/sweden_dist}
The values were picked arbitrarily and the value agreed upon to be used should be between 10 and 15 percent, as those values are the easiest to look at without losing sense of shape of the area. The performance tests still use the value of 20 percent to test the algorithms under less favourable conditions, since the larger the value of $d$, the denser and more time consuming to build the graph.

The tests use random inputs to benchmark the performance of the algorithms. Both uniform and clustered inputs are tested. The uniform inputs are a simple collection of $N$ points whose Cartesian coordinates are randomly chosen between $0$ and $100$. The clustered inputs are generated by choosing a random number of points $c$ between 10 to 20 points. Each of these points acts as a centre and is chosen by randomly generating two Cartesian coordinates between 0 and 100. Around each of these, $N/c$ points were generated by randomly choosing an angle $\Theta$ (between $0$ and $2\pi$) and a distance $\rho$ (between $10$ and $20$), which represent the polar coordinates around their respective central point. These tests generate circular clusters, with more density towards the centre.

Each test is performed 30 times, with shared seeds between the algorithms. The times listed do include the input scanning. All algorithms are implemented using ANSI C89 and compiled using \emph{gcc 5.1.0}. The machine and software specifications are listed in table \ref{tab:specs2}

\begin{table}[H]
	\begin{center}
		\begin{tabular}{|c|c|}
			\hline
			Operating System  & Arch Linux 3.14.4 x64\\\hline
			CPU & Intel i7 Dual-Core, 2GHz\\\hline
			Memory & 8 GB, 1600 MHz \\\hline
			Storage &  Solid-State Drive, 300 MB/s (read)\\\hline
		\end{tabular}
		\caption{Machine specifications}
		\label{tab:specs2}
	\end{center}
\end{table}

\input{Figures/kd_ls_t}

As Figure \ref{fig:kd_ls_t} shows, the line sweep method performs faster in both the uniform and the clustered data. The lower expected complexity of the \kdtrees does not compensate the larger overhead in their construction. It can also be seen that both algorithms become slower the larger the number of points, as was expected.

Applying both methods to any given case gives similar results. The coverage algorithm picks the points with most neighbours. Since two different points may have the same number of neighbours, and the two algorithms sort the points in two different ways, there may occur a case where the sets are completely different, based on which point is select first in case of a draw. The values still fall bellow the upper bound for the approximation factor, and no algorithm should have the advantage. Figure \ref{fig:kd_ls_k} shows the number of points selected by each algorithm in the same instances as above.

\input{Figures/kd_ls_k}

Figure \ref{fig:kd_ls_k} shows that no algorithm has the advantage on the number of points chosen. In fact, the results overlap each other almost completely. This suggests that the best algorithm to use is the line sweep algorithm, as it takes less time to compute similar results.

\section{Heuristic Speed-ups}
The approximation algorithm runs rather efficiently, taking under 1 second for 30 thousand points for the uniform inputs, and under 3 seconds for the clustered inputs. This result can be improved by employing some heuristic filtering methods to the inputs. However, by using these approaches, the guarantee of the quality is lost. Nevertheless, the results should still be considered for larger number of points.
\subsection{Sampling}
The simplest method to accelerating the algorithms is to simply ignore a random set of those points. With the smaller sample of points, the algorithm should run faster. If the points are removed uniformly, then the shape should still be kept.

Figure \ref{fig:ls_rs_t} shows the time difference between the regular line sweep and the sampled input.

\input{Figures/ls_rs_t}

The graph in Figure \ref{fig:ls_rs_t} proves that the sampling is indeed faster. However, upon closer inspection, it can be noted that the result is not optimal. Figure \ref{fig:ls_rs_k} shows that the number of selected points is inferior:

\input{Figures/ls_rs_k}

This can be explained by the fact that the algorithm is not accounting for all points. In fact, the number of cases where a point is left isolated in the full set algorithms is decreased by the factor of the sampling in these algorithms. This means that not all points are being covered, and the number of selected points decreases. Figure \ref{fig:ls_rs_greece} shows this effect in a real map:

\input{Figures/ls_rs_greece}

As shown, the sampled subset does not cover the most isolated points. In fact, the very last point to be covered in the original algorithm, the island in dark red at the westernmost point in the map, is not present in the sample, as as such, is not covered. This method is therefore not suitable for the initial requirements, as very isolated points have only a small chance of being represented.

\subsection{Two-phase filtering}
A solution to the random sampling algorithm problems is to perform two passes of the approximation algorithm for the geometric disk cover problem. The first pass over the points is done with a very small radius. This can be done very quickly, since the range search will only have to look in a very small area. Using a small distance accelerates the algorithm, as shown in Figure \ref{fig:ls_rs_t}. This means that many of the points that are close to each other will be discarded, but a representative neighbour will be left in its place. This means that isolated points will not be discarded. The resulting set will be much smaller than the original, whilst still keeping a representativeness degree. The second pass, with the final intended distance will take advantage of the much smaller set of points for the speed up. The \emph{CPU} times for two instances of the algorithm can be seen in Figure \ref{fig:ls_bp_t}. The values of the first pass distance $d^\prime$ are given as a percentage of the final pass distance $d$. One of the instances uses $d^\prime=5\%$ and the other $d^\prime=10\%$

\input{Figures/ls_bp_t}

As it can be seen, the two-phase filtering is very fast and does not seem to vary a lot with the total number of points. This is because the first pass eliminates a very large number of points, and the resulting set will be very similar for different values of $N$, since the intermediate subset is still representative of the original. Figure \ref{fig:ls_bp_k} shows the variation in size of the final subset chosen by the biphasic filtering:

\input{Figures/ls_bp_k}

The graph shows that despite the two-phase algorithms returning lower numbers of $K$, they are not as low as the random sampling. The resulting set does not necessarily cover all the points with disks of radii $d$. Because the first pass transforms the set into disks of radii $d^\prime$, and the second pass only considers their centre point as the measure of cover, then there can be points that are further away from the centroids than $d$, the maximum distance being $d+d^\prime$, as show in Figure \ref{fig:bp_error}:

\input{Figures/bp_error}

This result has some effect on the quality, but it is not nearly as noticeable as the results from the random sampling. Figure \ref{fig:bp_sweden} shows the compared output between the original line sweep algorithm, and the different versions of the two-phase algorithm, as well as their intermediate phases.

\input{Figures/bp_sweden}

This final result shows that the biphasic filter, especially with $d^\prime=0.1d$, can yield very approximate results at a fraction of the time, thus making a good candidate to use with inputs larger than the regular Line Sweep method can handle.